Supporting Information for ``Effects of pre-existing structures on the seismicity of the Charlevoix Seismic Zone"

### 2019-06-29 Rebuttal_Letter.tex

\begin{response}{4. Section 2.3: You did not mention the values of bulk modulus and shear modulus applied in your calculation.}

  PyLith, the modeling software used in this study, internally converts P and S wave velocities and density to bulk and shear moduli. For readers' reference, however, we now list representative values of elastic moduli in Table 1. 
  %\add[EC]{For readers' reference, however, we now list representative values of elastic moduli in Table 1.}\note[EC]{Seun, please take care of this.}
\end{response}

\begin{response}{5. Line 154-156: As I said in the main concern, you need to justify the reduction of 10\% from the velocity model to represent the impact crater area, instead of, for example, 2\%, 5\%, or 20\%.}

  The revised manuscript now includes a justification that the 10\% reduction of the density in the impact structure represents damage made to crustal rocks by the impact (Lines 181-186). Since it is not known quantitatively how much the impact affected the crustal density and elastic moduli, we ran extra models to explore the effects of density reduction from 0 to 20 \%. No significant changes in the model results were found in these models. These new results are included in the supplementary information and Fig. S3 therein.
\end{response}

@@

\begin{response}{7. Line 226-231: When evaluate the modeling result, I suggest the authors to quantify the fitting between the simulated stress change and the earthquake hypocenters. It will help convincing the readers of preferred model parameters. For example, I do not see it necessarily to prefer the frictional coefficient equals to 0.3 over 0.4 based on Figure 11.}

   We have added a quantitative assessment of the correlation between the change in differential stress and the observed seismicity by determining the percentage of hypocenters in the region with $\Delta\sigma_{D}^\prime \ge$ 0.5\% without accounting for uncertainty in hypocenter relocation (Line 205-209). This metric is now used for describing the model results and comparing different models throughout the revised manuscript. We also added a discussion on the choice of the friction coefficient in Lines 512-518.
    %: 1) the change in differential stress maps changes with fault geometry, and we have established an along-strike variation in the fault dips in this study; and 2) the 3D velocity model is very well defined in the upper 5 km depth which may introduced an uncertainty in the hypocenter relocation.
\end{response}

\begin{response}{8. Line 312: You need to justify the mesh size of 4 km in Section 2.1.}

We decided to run the models for various friction coefficients at the 4 km resolution to acquire necessary information without increasing computational cost too much. The low-friction models suffer slow convergence rate in the non-linear solution scheme. For instance, the models with $\mu$ equal to 0.1 and 0.2 were let run for 4 weeks on 40 cores to progress only to about half of the target model time. Running all the models in this group at the same mesh resolution of 4 km ensures intra-group consistency but at the same time, comparison with 2 km-resolution models is also justified because the descritization error does not significantly increase as the resolution increases from 2 to 4 km (Fig. S1). This discussion is now included in Lines 118 to 127 of the revised manuscript.
\end{response}

### 2019-06-29 Supporting_Information.tex
saving progress.

%IMPORTANT NOTE ON FIGURES AND TABLES
% Placeholders for figures and tables appear after the \end{article} command, after references.
% DO NOT USE \psfrag or \subfigure commands.
%
\usepackage{rotating}
\usepackage{graphicx}

@@

From $\Delta$ABC in Fig.~\ref{S2},
\[
 \sin \alpha = \frac{\sigma_{1}-\sigma_{3}}{2} / \left[\frac{250}{3} + \frac{\sigma_{1}+\sigma_{3}}{2} \right],
\]
%
\[
 \sin \alpha = \frac{3(\sigma_{1}-\sigma_{3})}{3(\sigma_{1}+\sigma_{3})+500},
\]
%
\[
 \sigma_{1} = \frac{\sigma_{3}+[\sigma_{3}+\frac{500}{3}] \sin(\alpha)}{1-\sin \alpha},
\]

Based on Byerlee's law, $\tau$  = 50 MPa + 0.6$\,\sigma_n$, under the assumption of absence of pore-fluid pressure along the rift faults, $\tan \alpha = 0.6$, hence, $\sin \alpha = 0.5145$. $\sigma_{3}$ corresponds to the lithostatic stress, and assuming it is equal to 250 MPa at 10 km depth, gives $\sigma_{1}$ = 956.4882 MPa. Hence, the differential stress at 10 km depth with no pore-fluid pressure ($\sigma_{D} = \sigma_{1} - \sigma_{3})$ is 706.4882 MPa.

@@

The percentage change in differential stress ($\Delta\sigma_{D}^{\prime}$), SH$_{\max}$ orientation ($\phi_{\max}$), and the change in $\phi_{\max}$ ($\Delta\phi_{\max}$) relative to the reference model (SNFR25) for models with different impact structure-crust density ratio (Fig. \ref{S3}).  The maps show no significant change with density ratio. We use density ratio of 10\% in this study to model a realistic damaged crustal rock. 
\vspace{10mm} %5mm vertical space

\begin{figure}[ht]
\setfigurenum{S3}
\includegraphics[width=25pc]{Figures/effect_of_density.png}
\caption{Effect of impact structure-crust density ratio on the $\Delta\sigma_{D}^{\prime}$, $\phi_{\max}$ and $\Delta \phi_{\max}^{\prime}$ at 10 km depth. Earthquakes within 2 km of each depth slice are represented by black dots in \ref{S3}A. The outline of the impact structure is represented as black circle. Each figure is 70 $\times$ 70 km.}
\label{S3}
\end{figure}

@@

The percentage change in differential stress ($\Delta\sigma_{D}^{\prime}$), SH$_{\max}$ orientation ($\phi_{\max}$), and the change in $\phi_{\max}$ ($\Delta\phi_{\max}$) relative to the reference model (SNFR25) at various depths from the SD70R100 model (Fig. \ref{S4}).  $\Delta\sigma_{D}^{\prime}$ shows increased $\sigma_{D}$ relative to SNFR25 throughout the entire length of the rift faults at and below 15 km depth, hence the model predicts earthquakes in those regions.

@@

\begin{figure}[ht]
\setfigurenum{S4}
\includegraphics[width=25pc]{Figures/SD70R100_S1.png}
\caption{Percentage change in differential stress ($\Delta\sigma_{D}^{\prime}$), SH$_{\max}$ orientation ($\phi_{\max}$), and the change in $\phi_{\max}$ ($\Delta\phi_{\max}$) relative to the SNFR25 at various depths depths from the SD70R100 model. Earthquakes within 2 km of each depth slice are represented by black dots in \ref{S3}A. Each figure is 70 $\times$ 70 km.}
\label{S4}
\end{figure}

@@

We test the response of the model results with different impact structure-crust moduli ratio (Fig. \ref{S5}). The model with a moduli ratio of 1.0 does not have the impact crater. The models show a better spatial correlation with the observed seismicity when the moduli ratio is 0.25.

@@

We compute $\Delta \sigma_{D}^{\prime}$, $\phi_{\max}$ and $\Delta \phi_{\max}^{\prime}$ in a series of models in which the rift faults dip at 70$^{\circ}$ (i.e., SD70R25) but their cohesion ($C$) is varied systematically at a constant coefficient of friction ($\mu$) of 0.3. We assume that the fault parameters are the same for each rift fault and uniform throughout the fault surfaces. Six models with 0, 3, 5, 10, 20 and 30 MPa are constructed (Fig. \ref{S6}).

Magnitudes of $\Delta\sigma_D^{\prime}$ and $\phi_{\max}$ decrease with increasing cohesion. $\Delta\sigma_D^{\prime}$ maps spatially overlap with the distribution of seismicity at all the fault cohesions, especially well when the cohesion is less than 5 MPa (Fig. \ref{S6}A). However, the models could not explain some earthquakes in the SW corner of the crater. The $\Delta\sigma_D^{\prime}$ is not affected by a change in cohesion between 0 and 5 MPa. However, $\phi_{\max}$ approaches the reference no-fault model as the cohesion increases (Fig. \ref{S6}B). As cohesion increases, $\Delta\phi_{\max}$ decreases in value and the rotation is concentrated in the fault region outside the crater (Fig. \ref{S6}C). $\Delta\phi_{\max}$ is up to 10.8$^\circ$ when the cohesion is less than 5 MPa.

Figure \ref{S6} shows that $\mu$ is an essential fault parameter needed to explain the seismicity of the CSZ because the value of $\mu$ affects the slope of the Mohr-Coulomb failure envelope. We use a cohesion value of 0 MPa to explain the spatial distribution of observed seismicity in the CSZ.

>>>>

%IMPORTANT NOTE ON FIGURES AND TABLES
% Placeholders for figures and tables appear after the \end{article} command, after references.
% DO NOT USE \psfrag or \subfigure commands.
%
\usepackage{rotating}
\usepackage[inline]{trackchanges}
\addeditor{EC}
\addeditor{OF}
\usepackage{graphicx}

@@

From $\Delta$ABC in Fig.~\ref{S2}, the illustration of the Mohr circle and the Byerlees law, we get
\[
 \sin \alpha = \frac{\sigma_{1}-\sigma_{3}}{2} / \left[\frac{250}{3} + \frac{\sigma_{1}+\sigma_{3}}{2} \right].
\]
%
%\[
% \sin \alpha = \frac{3(\sigma_{1}-\sigma_{3})}{3(\sigma_{1}+\sigma_{3})+500},
%\]
%
After simpligying and solving for $\sigma_{1}$, we ge the following expression:
\[
 \sigma_{1} = \frac{\sigma_{3}+[\sigma_{3}+\frac{500}{3}] \sin(\alpha)}{1-\sin \alpha}.
\]

Based on Byerlee's law, $\tau$  = 50 MPa + 0.6$\,\sigma_n$, under the assumption of absence of pore-fluid pressure along the rift faults, $\tan \alpha = 0.6$, hence, $\sin \alpha = 0.5145$. $\sigma_{3}$ corresponds to the lithostatic stress amounting to 250.0 MPa at 10 km depth \add[EC]{for a density of XXXX kg/m$^{3}$}. With these values and the above expression, we get $\sigma_{1}$ = \annote[EC]{956.5}{Pay attention to the number of significant figures! I don't think our calculations can afford 7 or 8-digit precision so I rounded off the original value leaving 4 digits.} MPa. The differential stress at 10 km depth with no pore-fluid pressure, defined as $\sigma_{D} = \sigma_{1} - \sigma_{3})$, is 706.5 MPa.

@@

\note[EC]{Turn the first long phrase into a sentence. Right now it reads like a figure caption but it shouldn't in a normal paragraph.} The percentage change in differential stress ($\Delta\sigma_{D}^{\prime}$), SH$_{\max}$ orientation ($\phi_{\max}$), and the change in $\phi_{\max}$ ($\Delta\phi_{\max}$) relative to the reference model (SNFR25) for models with different impact structure-crust density ratio (Fig. \ref{S3}).  The maps show no significant change with density ratio. In this study, we choose the density ratio of 10\% to represent damaged crustal rock. 
\vspace{10mm} %5mm vertical space

\begin{figure}[ht]
\setfigurenum{S3}
\includegraphics[width=25pc]{Figures/effect_of_density.png}
\caption{\note[EC]{Symbol descriptions are missing. Include those in Fig. S4 caption.} Effect of impact structure-crust density ratio on the $\Delta\sigma_{D}^{\prime}$, $\phi_{\max}$ and $\Delta \phi_{\max}^{\prime}$ at 10 km depth. Earthquakes within 2 km of each depth slice are represented by black dots in \ref{S3}A. The outline of the impact structure is represented as black circle. Each figure is 70 $\times$ 70 km.}
\label{S3}
\end{figure}

@@

\note[EC]{Turn the first long phrase into a sentence. Right now it reads like a figure caption but it shouldn't in a normal paragraph.} The percentage change in differential stress ($\Delta\sigma_{D}^{\prime}$), SH$_{\max}$ orientation ($\phi_{\max}$), and the change in $\phi_{\max}$ ($\Delta\phi_{\max}$) relative to the reference model (SNFR25) at various depths from the SD70R100 model (Fig. \ref{S4}).  $\Delta\sigma_{D}^{\prime}$ shows increased $\sigma_{D}$ relative to SNFR25 throughout the entire length of the rift faults at and below 15 km depth, hence the model predicts earthquakes in those regions.

@@

\begin{figure}[ht]
\setfigurenum{S4}
\includegraphics[width=25pc]{Figures/SD70R100_S1.png}
\caption{Percentage change in differential stress ($\Delta\sigma_{D}^{\prime}$), SH$_{\max}$ orientation ($\phi_{\max}$), and the change in $\phi_{\max}$ ($\Delta\phi_{\max}$) in SD70R100 relative to SNFR25 at various depths. Earthquakes within 2 km of each depth slice are represented by black dots in \ref{S3}A. Each figure is 70 $\times$ 70 km.}
\label{S4}
\end{figure}

@@

We run 4 models in total to assess the effects of impact structure-to-crust moduli ratio assuming that the impact structure was softened. The ratio equal to 1, for instance, means that the impact structure was not softened and thus has the elastic moduli with those of the surrounding crust. The models show a better spatial correlation with the observed seismicity when the moduli ratio is 0.25.

@@

We compute $\Delta \sigma_{D}^{\prime}$, $\phi_{\max}$ and $\Delta \phi_{\max}^{\prime}$ in a series of models in which the rift faults dip at 70$^{\circ}$ (i.e., SD70R25) but their cohesion ($C$) is varied from 0 to 30 MPa while the friction coefficient ($\mu$) is fixed at 0.3 (Fig.~\ref{S6}). We assume that the fault parameters are the same for each of the three rift faults and uniform throughout the fault surfaces. %Six models with 0, 3, 5, 10, 20 and 30 MPa are constructed (Fig. \ref{S6}).

Magnitudes of $\Delta\sigma_D^{\prime}$ and \annote[EC]{$\phi_{\max}$}{Fig. S6B doesn't support this.} decrease with increasing cohesion. $\Delta\sigma_D^{\prime}$ maps spatially overlap with the distribution of seismicity for all the fault cohesions\annote[EC]{, especially well when the cohesion is less than 5 MPa}{What makes you say so? Can't you use the new metric for this argument?} (Fig. \ref{S6}A). However, the models could not explain a few earthquakes in the SW corner of the crater, where $\Delta\sigma_D^{\prime}$ is negative. The $\Delta\sigma_D^{\prime}$ is not affected by a change in cohesion between 0 and 5 MPa. However, $\phi_{\max}$ approaches the reference no-fault model as the cohesion increases (Fig. \ref{S6}B). As cohesion increases, $\Delta\phi_{\max}$ decreases in value and the rotation is concentrated in the fault region outside the crater (Fig. \ref{S6}C). $\Delta\phi_{\max}$ is up to 10.8$^\circ$ when the cohesion is less than 5 MPa.

\note[EC]{Why do you make remarks on $\mu$ mentioning cohesion only little in the paragraph for the effects of cohesion?} Figure \ref{S6} suggests that $\mu$ is an essential fault parameter needed to explain the seismicity of the CSZ because the value of $\mu$ affects the slope of the Mohr-Coulomb failure envelope. We present models with zero cohesion in the main text because to explain the spatial distribution of observed seismicity in the CSZ.
